---
title: "The STAR-ting Point for STAR and EXOPLANET Exploration!"
author: "KelliAnn Ramirez, Salley Fang, Claire Zhang"
date: "11/12/2022"
output: html_document
---

# CHONK!
```{r}
# TO DO:
# 1. create a new .R file that calculates 5 diff values (summary info)
#    - the file is summary_values.R, has everything
# 2. Write a dynamic paragraph abt summary info in index.Rmd
# 3. Create table for summary info (?)
# 4. Create chart/values for fires.csv
#    - charts present in fire_charts.R
# 5. Update project proposal (edit project goals)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Dynamic Paragraph
The values we chose display crucial information to help answer our research questions. The ***first two values*** were taken from `earth-land-temps.csv`. **Value 1** (`highest_month_value`) includes the month and value of the highest temperature change across all years and countries recorded: `March: 0.59834155983724`. **Value 2** (`lowest_month_val`) includes the month and value of the lowest temperature change across all years and countries recorded: `December: 0.430323834725771`. These values show the narrow range of temperature increases throughout the year and when higher temperature increases occurred to find connections between temperature increase and time of year. The ***next three values*** were taken from `exoplanets.csv`. **Value 3** (`percent_bigger`) is the percent of exoplanets bigger than the Earth (95.39%) and **Value 4** (`percent_smaller`) is the percent smaller than Earth (4.61%). These percentages are important in learning about the features that can make exoplanets habitable and giving an idea of the likelihood of finding them. **Value 5** (`percent_habitable`) is the chance we will find a habitable exoplanet given the ones in the dataset: 0.81%. This gives viewers a better understanding of the difficulty in finding other habitable planets. The ***last two values*** were taken from `fires.csv`. **Value 6** (`fire_diff`) is the difference between the number of fires in the earliest recorded year (1992) and latest (2015): 30,811. **Value 7** (`percent_diff`) is the percent difference between fires from the earliest and latest years (1992-2015): 99.3791%. These values are important as they clearly reflect the dramatic increase in fires in just over thirty years. The numerical increase and percentages both reflect how dramatic the change is in different ways.

## Summary Information Table
```{r, warning=F, echo=F}
library(knitr)
source("../source/summary_values.R")
kable(summary_values_df)
```

### Sourcing Oceania Temperature Changes 
```{r, echo = FALSE, out.width='50%'}
source("../source/Oceania_Temp_Change_Chart2.R")
oceania_bar_graph
oceania_avg_temp <- avg_oceania_temp_change

```
  
This bar graph is an example of the types of data visualizations we will include on our final website. It displays the _average temperature change_ across all months and years included in the dataframe (Jan-Dec, 1961-2019) for each _country_ in the selected continent. This average change is displayed in a bar graph where the ***x-axis*** is the _country’s name_ (`Area`) and the ***y-axis*** shows the _average change in temperature_ (`avg_change`). As seen in _Oceania Temperature Change_, all countries (besides Nauru) showed increases in temperature that averaged `0.3345707` degrees. 

We want our final website to allow users to filter data by continents so it displays each country in that continent’s average temperature change. This can reveal how the changes in temperature differ by region, as different continents likely have differing patterns in temperature increase. We decided to include this feature because we wanted users to see a more ‘zoomed in’ view of temperature. Rather than seeing data on a large scale, we wanted to display it by continents so viewers can easily spot patterns and see temperature changes under different parts of the world. Since different climates likely correspond to differing changes in temperature, our visualization can display unique changes by area clearer.

### Sourcing Habitable Exoplanets 
```{r, echo = FALSE, out.width='50%'}
source("../source/exoplanet-chart-code.R")
bar_planet_temp_count
habitable_exoplanets
```

It is important when looking for possible homes beyond Earth that we make sure we can live on said planet. One of the most important things humans need is for the planet to have a comfortable climate. The chart above on the left reveals to us that most of the known exoplanets out there are roughly between 500 K and 750 K (~440 F - 890 F), which unfortunately means that most of the exoplanets out there are uninhabitable for humans. Humans are comfortable in temperatures from about 273 K to 300 K (~33 F - 80 F) and there are only roughly 41 possibly habitable exoplanets by temperature out of the 5044 known planets in this dataset. This is only a 0.81% chance that the exoplanet we arrive at is the right temperature. The chart above on the right shows planets at comfortable temperatures for humans. When compared to the left chart, this chart reveals to us how little habitable planets we know of in the context of all the exoplanets we’ve discovered and even the entire universe. This is important to include because we need to be aware that not every exoplanet out there can sustain human life. There are many factors that affect a planet that we may not know about until we get there such as we can’t land on the surface for some reason or the weather on the planet surface is so bad that it’s intolerable.

### Sourcing Fires by State Map
```{r, echo=F}
# fires.csv chart reference code here
source("../source/fires_charts.R")
year_plot
map_1998
map_2015

```

Paragraph answering: Beneath each chart, you must include a brief paragraph, describing 
**why** you included the chart and **what** information it reveals, including notable observations
and insights from the chart.
