---
title: "The STAR-ting Point for STAR and EXOPLANET Exploration!"
author: "KelliAnn Ramirez, Salley Fang, Claire Zhang"
date: "11/12/2022"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Dynamic Paragraph
The values we chose display crucial information to help answer our research questions. The ***first two values*** were taken from `earth-land-temps.csv`. **Value 1** (`highest_month_value`) includes the month and value of the highest temperature change across all years and countries recorded: `March: 0.6`. **Value 2** (`lowest_month_val`) includes the month and value of the lowest temperature change across all years and countries recorded: `December: 0.43`. These values show the narrow range of temperature increases throughout the year and when higher temperature increases occurred to find connections between temperature increase and time of year. The ***next three values*** were taken from `exoplanets.csv`. **Value 3** (`percent_bigger`) is the percent of exoplanets bigger than the Earth (95.39%) and **Value 4** (`percent_smaller`) is the percent smaller than Earth (4.61%). These percentages are important in learning about the features that can make exoplanets habitable and giving an idea of the likelihood of finding them. **Value 5** (`percent_habitable`) is the chance we will find a habitable exoplanet given the ones in the dataset: 0.81%. This gives viewers a better understanding of the difficulty in finding other habitable planets. The ***last two values*** were taken from `fires.csv`. **Value 6** (`fire_diff`) is the difference between the number of fires in the earliest recorded year (1992) and latest (2015): 30,811. **Value 7** (`percent_diff`) is the percent difference between fires from the earliest and latest years (1992-2015): 99.38%. These values are important as they clearly reflect the dramatic increase in fires in just over thirty years. The numerical increase and percentages both reflect how dramatic the change is in different ways.

## Summary Information Tables
### `earth_temp_summary_table`
```{r, echo=F}
library(knitr)
source("../source/summary_values_table.R")
kable(earth_temp_summary_table)
```
This table displays the ***average change in temperature*** of _all countries in the dataframe across all years and months_ that were counted. We wanted to include this table because it shows how every country (besides _Nauru_ and _Pitcairn Islands_) showed increases in temperature. It tells us how temperature increases are not a trivial or randomly occurring change but a worldwide one that should be taken seriously.

### `exoplanets_summary_table`
```{r, echo=F}
source("../source/summary_values_table.R")
library(knitr)
kable(exoplanets_summary_table)
```
This table displays the ***average orbital day*** for exoplanets with an _orbital period between 200 to 700 days_ for each applicable ***discovery method***, not including ones that were used to discover exoplanets with orbital periods not in the range (200-700). We wanted to include this table to give a better idea as to the discovery methods that can most accurately find the exoplanets with orbital periods similar to Earth's. This tells us that ***Eclipse Timing Variations*** has found exoplanets with an average orbital period most similar to Earth. Astronomers may want to utilize this method over others in discovering more exoplanets.

### `fires_summary_table`
```{r, warning=F, echo=F}
library(knitr)
source("../source/summary_values_table.R")
kable(fires_summary_table)
```
This table displays the ***average fire size*** (in acres) for _all years_ included in the dataframe. We wanted to include this table to show how much the size of fires has increased in just 23 years. In **1992**, the average fire size was ***12.01 acres***. By 2015, the average fire size had increased to ***110.31 acres***. This dramatic increase reflects how much of a fast growing issue forest fires are becoming in recent years.

## Summary Charts
### Sourcing Oceania Temperature Changes 
```{r, echo = FALSE, out.width='50%'}
source("../source/Oceania_Temp_Change_Chart2.R")
plot(temp_change_country$avg_change)
plot(yearly_avg_change$avg_change_total)
oceania_bar_graph
```
  
This bar graph is an example of the types of data visualizations we will include on our final website. It displays the _average temperature change_ across all months and years included in the dataframe (Jan-Dec, 1961-2019) for each _country_ in the selected continent. This average change is displayed in a bar graph where the ***x-axis*** is the _country’s name_ (`Area`) and the ***y-axis*** shows the _average change in temperature_ (`avg_change`). As seen in _Oceania Temperature Change_, all countries (besides Nauru) showed increases in temperature that averaged `0.33` degrees. 

We want our final website to allow users to filter data by continents so it displays each country in that continent’s average temperature change. This can reveal how the changes in temperature differ by region, as different continents likely have differing patterns in temperature increase. We decided to include this feature because we wanted users to see a more ‘zoomed in’ view of temperature. Rather than seeing data on a large scale, we wanted to display it by continents so viewers can easily spot patterns and see temperature changes under different parts of the world. Since different climates likely correspond to differing changes in temperature, our visualization can display unique changes by area clearer.

### Sourcing Habitable Exoplanets 
```{r, echo = FALSE, out.width='50%'}
source("../source/exoplanet-chart-code.R")
bar_planet_temp_count
habitable_exoplanets
```

It is important when looking for possible homes beyond Earth that we make sure we can live on said planet. One of the most important things humans need is for the planet to have a comfortable climate. The chart above on the left reveals to us that most of the known exoplanets out there are roughly between _500 K and 750 K_ (~440 F - 890 F), which unfortunately means that most of the exoplanets out there are uninhabitable for humans. Humans are comfortable in temperatures from about _273 K to 300 K_ (~33 F - 80 F) and there are only roughly _41 possibly habitable exoplanets_ by temperature out of the 5044 known planets in this dataset. This is only a **0.81%** chance that the exoplanet we arrive at is the right temperature. The chart above on the right shows planets at comfortable temperatures for humans. When compared to the left chart, this chart reveals to us how little habitable planets we know of in the context of all the exoplanets we’ve discovered and even the entire universe. This is important to include because we need to be aware that not every exoplanet out there can sustain human life. There are many factors that affect a planet that we may not know about until we get there such as we can’t land on the surface for some reason or the weather on the planet surface is so bad that it’s intolerable.

### Sourcing Fires by State Map
```{r, echo=F, out.width='50%'}
# fires.csv chart reference code here
source("../source/fires_charts.R")
year_plot
map_1998
map_2015

```

The `year_plot` bar graph is important in seeing the gradual increase in fires over the course of
23 years (1992-2015). It shows how dramatic the number of fires has jumped in just the previous five
years from 2010 to 2015. The increase in fires from 1992 to 2010 revealed a more gradual increase,
reflected in the shift from lighter oranges to darker yellows to darker oranges. But after 2010,
that color shifted suddenly to a deep red.

The `map_1998` choropleth map was included to give viewers a map to compare the more recent fires map
(`map_2015`) to. The parts of the map shaded in gray represent states in which fire counts were not
given. The map reveals the distribution of fires throughout the country. The distribution of fires
was pretty uniform, besides in Mississippi. 

The `map_2015` map was included to show how dramatic the increase in fires has been. Compared to the
`map_1998` fire distribution, this map shows dramatic changes in certain states, especially ***California,
Kansas, and New York***. In the 1998 map, these states either had an unreported number of fires (_Kansas_, 
_New York_) or far fewer than in 2015 (_California_). This jump was not over the course of centuries or
multiple decades, but over the course of just over 20 years. This can reveal how urgent of an issue fires
has become in recent years and the importance of dealing with climate change.

